Actuator line modeling of vertical-axis turbines

### 2016-05-04 paper.tex
Do not speculate as to why DMST models do poorly for high solidity; resolves #18

Very fast solution times can be obtained with blade element momentum (BEM)
models, such as the double multiple streamtube (DMST) approached developed by
Paraschivoiu~\cite{Para1988}. However, their poor performance for high loading
or streamwise induction factors, common for rotors with high
solidity~\cite{Joo2015} $Nc/R$ (and/or chord-to-radius ratio $c/R$), where $N$
is the number of blades, $c$ is the chord length, and $R$ is the maximum rotor
radius, makes them less attractive, particular for marine hydrokinetic
applications where rotors of the same size will encounter approximately an order
of magnitude higher torque compared to wind turbines. Their description of the
flow field is very crude as well, simply computing the momentum contained within
discretized streamtubes, which cannot resolve any of the complicated flow
structures shed by the rotor, limiting their applicability for array analysis.

@@

    \caption{Vector diagram of velocity and forcing on a cross-flow turbine
        blade element. Note that the free stream velocity $U_\infty$ is from top to
        bottom (identical to $U_in$ for purely geometric calculations), the blade
        chord (dash-dotted line) is coincident with the tangential velocity (i.e.,
        zero preset pitch, which would offset the geometric angle of attack
        $\alpha$), and the drag vector is magnified by a factor of two
        (approximately, relative to the lift vector) to enhance visibility.}

>>>>

Very fast solution times can be obtained with blade element momentum (BEM)
models, such as the double multiple streamtube (DMST) approached developed by
Paraschivoiu~\cite{Para1988}. However, their poor performance for rotors with
high solidity~\cite{Joo2015} $Nc/R$ (and/or chord-to-radius ratio $c/R$), where
$N$ is the number of blades, $c$ is the chord length, and $R$ is the maximum
rotor radius, makes them less attractive, particular for marine hydrokinetic
applications where rotors of the same size will encounter approximately an order
of magnitude higher torque compared to wind turbines. Their description of the
flow field is very crude as well, simply computing the momentum contained within
discretized streamtubes, which cannot resolve any of the complicated flow
structures shed by the rotor, limiting their applicability for array analysis.

@@

    \caption{Vector diagram of velocity and forcing on a cross-flow turbine
        blade element. Note that the free stream velocity $U_\infty$ is oriented
        from top to bottom (identical to $U_in$ for purely geometric calculations),
        the blade chord (dash-dotted line) is coincident with the tangential
        velocity (i.e., zero preset pitch, which would offset the geometric angle of
        attack $\alpha$), and the drag vector is magnified by a factor of two
        (approximately, relative to the lift vector) to enhance visibility.}
